### 2018-10-04 Workflow.tex
update of paper files

\documentclass{article}
\usepackage{hyperref}
\usepackage[dvipsnames]{color}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}
\title{Workflow for CLM5 paper}

Scripts for these analyses should all be found in the public github repo at:\\
{\color{blue}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS}}}\\

The scripts are specific to usage on the NCAR Cheyenne machine...\\

{\color{blue} Scripts, on github are in blue}\\
{\color{cyan} Paths on scratch are in cyan}\\
{\color{red}  .nc files are in red}\\
{\color{magenta} Figures on github are in magenta}\\

\section{Obtain CLM5 code}
cd /glade/p/work/rfisher/git/\\
(this is 'gitpath' from now on)

git clone https://github.com/ESCOMP/ctsm 
\section{Parameters}

\subsection{Make new ensemble param files with high leaf allocation }

Use PFT script

{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/FACE_clm5_ens_oaat.m}}}\\ 
commit b6ba94

Using the basefile. \\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/clm5_params.c171017_highleaf.nc}}} \\
commit c539e89

Output location..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi11.nc }}} \\
commit 54f863b

Copy default file into '00'..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi00.nc }}} \\
commit 61517c5

\section{Spin-up}
\subsection{Make 1850 spin-up default runs from scratch}
Make single point case from scratch for each site by running this script (based on Charlie's FATES version) three times once for each site (that could be improved on!)..:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/fromscratch_CTSM_make_site_spinup_case_defaults}}}\\
commit 38ced1

The resulting cases are \\
{\color{cyan}{\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_US-ORN}}}\\
\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.bci_0.1x0.1_v4.0i}}\\
\small{\url{/glade/scratch/rfisher/CLM5_SPR_SPIN_HL_.1x1pt_Br-cax}}}\\

These are all run out to year 321.

\subsection{Make 1850 spin-up ensembles for each site}

Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
commit 2bb89c5

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI**}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PI**}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PI*}}}\\

These are all run to y201.

\subsection{Make 1850 spin-up defaults for each site}

Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults}}}\\
commit 0d2894a

Case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_PI00}}}

\subsection{Analyse output of spinup cases}

Run analysis script on spinup ensemble output (at 1850 CO2). \\
{\color{blue}\small{\url{/glade/p/work/rfisher/git/CLM5_FACE_EXPTS/matlab/CLM5_FACE_senstest_output_mar18.m}}}\\
commit 7043e8b

Output here\\
{\color{magenta}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS/tree/master/matlab/figures\frac_deviation_p1CLM5_1x1pt_Br-cax_ens_PI_y1.png}}}\\
commit 1f9a57c

\section{Transients}

\subsection{Make transient defaults}
Use this script. It employs the 20tr default CO2 ramping, but with an 1850 climate. .\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults}}}\\
commit 36ced2a

Case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI00}}}

\subsection{Make transient ensemble}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}}}\\
commit 7b8bca7

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_PI**}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_PI**}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_PI**}}}

\subsection{Analyse transient ensemble}

\subsection{Elevated CO2}
\subsection{Make elevated CO2 file}
Use this script\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers.m}}}\\
commit 0d4fb93

Which generates\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
commit 0d4fb93

\subsection{Make elevated CO2 defaults}
Use this script:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
commit 8e6a3a5

Which uses these user$\_$nl$\_$datm files:\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_Br-cax}}}\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_bci_0.1x0.1_v4.0i}}}\\
{\color{blue}\small{\url{gitpath/co2_files/user_nl_datm_FACE_transient_ELEV_1x1pt_US-ORN}}}\\
commit 4cb99a4

And this user$\_$datm.streams.txt.co2tseries.20tr file\\
{\color{blue}\small{\url{gitpath/co2_files/user_datm.streams.txt.co2tseries.20tr}}}\\
commit 41d4aeb

Which references the previous co2 file\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/co2_datm_ORNL_ELEV.nc}}}\\
commit 0d4fb93

The case names are:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI00}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI00}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI00}}}

\subsection{Make elevated CO2 ensemble}
Use this script:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
commit ffe1b6b

The case names are like:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI**}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI**}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI**}}}

\subsection{Analyze elevated CO2 runs}
Use this script:\\
{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
commit ec202a0

Creates these figures:
{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
commit 23ce56e

\section{Make runs with LUNA off}

\subsection{Make ensemble spinup runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
commit 88b6508

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI15*}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PI15*}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PI15*}}}\\

\subsection{Make default spinup runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_defaults_lunaoff}} }\\
commit 88b6508

Case names are along the lines of:\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_PI00_noluna}}\\
\small{\url{CLM5_1x1pt_Br-cax_ens_PII00_noluna}}\\
\small{\url{CLM5_1x1pt_US-ORN_ens_PII00_noluna}}}\\

\subsection{Make transient runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ensembles}} }\\
commit acdee08

The case names are like:\\
{\color{cyan}\small{\url{CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI15*}}}\\
{\color{cyan}\small{\url{CLM5_bci_0.1x0.1_v4.0i_ens_transient_ELEV_PI15*}}}\\
{\color{cyan}\small{\url{CLM5_1x1pt_US-ORN_ens_transient_ELEV_PI15*}}}

\subsection{Make default transient runs with LUNA off}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_defaults_lunaoff}} }\\
commit b26f487

\section{Correct the CO2 fertilization to be all year}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/modify_co2_drivers_allyear.m }} }\\
commit 723bbf3

File is here: 
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/mco2_datm_ORNL_ELEV_allyear.nc }} }\\
commit 8604dd9

\subsection{Restart ensemble transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_ensembles}}}\\
commit 9f8c771

\subsection{Restart default transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults}}}\\
commit cc6e00e

\subsection{Restart LUNA defaults transients with new CO2 file}
Use.\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_transient_ELEV_defaults_lunaoff}}}\\
commit cc6e00e

\subsection{Analyze new elevated CO2 runs}
Use this script:\\
{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_apr18.m}}}\\
commit  7eab468

Creates these figures:
{\color{magenta}\small{\url{gitpath/matlab/figures/frac_deviation_CO2_response_1CLM5_1x1pt_Br-cax_ens_transient_ELEV_PI_y1.png}}}\\
commit 23ce56e

\section{Add parameters to look at open/closed N cycle}
\subsection{Make namelist files witht this matlab script}
{\color{blue}\small{\url{gitpath/matlab/make_namelist_files.m}}}\\
commit 532c208

Files are in here:
{\color{magenta}\small{\url{gitpath/matlab/namelists/**}}}\\
commit c7633e9

\subsection{Make spinup runs for namelist parameters}
{\color{blue}\small{\url{gitpath//case_scripts/make_site_spinup_ensembles}}}\\
commit faa49a9

\subsection{Analyze spinup results for namelist parameters} 
matlab script modified
{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_mar18.m
}}}\\
e04d71

\subsection{Make transient runs for namelist parameters}

{\color{blue}\small{\url{gitpath//case_scripts/case_scripts/make_site_transient_ELEV_ensembles}}}\\
commit 96fafee

\subsection{Analyze transient runs for namelist parameters}

{\color{blue}\small{\url{gitpath/matlab/CLM5_FACE_senstest_output_ELEV_mar18.m}}}\\
commit fe1a59

\section{Modify range of FlexCN parameters to completely stop C:N flexibility?}

\subsection{Modified ranges of flex\_cn\_c parameters}

{\color{blue}\small{\url{gitpath/matlab/FACE_clm5_ens_oaat.m 
}}}\\
commit b3f67e9

Parameter files here:
Output location..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_highleaf_pi13*.nc }}} \\
commit 823385e

\subsection{Set up ensemble runs just for parameter 13}
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/make_site_spinup_ensembles}} }\\
b3f67e9

\subsection{Set up transient runs just for parameter 13}
{\color{blue}\small{\url{gitpath//case_scripts/case_scripts/make_site_transient_ELEV_ensembles}}}\\
b3f67e9

\subsection{Analyze new runs}

\section{Add Nitrogen fertilization experiments}
\subsection{Made new +5gCm2 file}

\subsection{Make transient NDEP ensembles}
{\color{blue}\small{\url{gitpath/case_scripts/make_site_transient_NDEP_ensembles}}}\\ 
commit 2e395ef

\subsection{Make transient NDEP defaults}
{\color{blue}\small{\url{gitpath/case_scripts/make_site_transient_NDEP_defaults}}}\\ 
commit 2e395ef

\section{Add additional sites}
added Barrow

\section{add temperature experiments}

\section{add reduced form global sensitivity analysis}

 \section{Figures}
 \subsection{Make Will figure}
 \subsection{Make initial LAI figure}
 \subsection{Make 3D figures of parameter space}

\section{Investigate differences in driving data for single sites vs reanalysis}

\end{document}