### 2018-10-29 workflow2.tex
Updates from ShareLaTeX

\documentclass{article}
\usepackage{hyperref}
\usepackage[dvipsnames]{color}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}
\title{Final Workflow for CLM5 paper}

Scripts for these analyses should all be found in the public github repo at:\\
{\color{blue}\small{\url{https://github.com/rosiealice/CLM5_FACE_EXPTS}}}\\

The scripts are specific to usage on the NCAR Cheyenne machine...\\

{\color{blue} Scripts, on github are in blue}\\
{\color{cyan} Paths on scratch are in cyan}\\
{\color{red}  .nc files are in red}\\
{\color{magenta} Figures on github are in magenta}\\

\section{Obtain CLM5 code}
cd /glade/p/work/rfisher/git/\\
(this is 'gitpath' from now on)

git clone https://github.com/ESCOMP/ctsm 
\section{Parameters}

Make ensemble parameter files using

{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/FACE_clm5_ens_oaat.m}}}\\

Using the basefile. \\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/clm5_params.c171017.nc}}} \\

Output location..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_defpft_pi11.nc }}} \\

Copy default file into '00'..\\
{\color{red}\small{\url{gitpath/CLM5_FACE_EXPTS/matlab/FACE_pftfiles/FACE_CLM_params_may1_defpft_pi00.nc }}} \\

\section{Spin-up}
\subsection{Make 1850 spin-up default runs from scratch}
Make single point case from scratch for each site by running this script (based on Charlie's FATES version) 'n' times once for each site (that could be improved on!)..:\\
{\color{blue}\small{\url{gitpath/CLM5_FACE_EXPTS/case_scripts/fromscratch_CTSM_make_site_spinup_case_defaults}}}\\

\end{document}