### 2019-07-05 referee_responses.tex

\documentclass{article}
\usepackage{hyperref}
\usepackage[dvipsnames]{color}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}

\textsf{Reviewer 1}

To be fully transparent I should declare that this review is written by Martin De Kauwe. I agreed to review the paper before I realised Anthony Walker was one of the authors (a potential conflict of interest). I subsequently informed the editor but they were keen for me to still review the manuscript.

\textsf{AR: Fair enough.}

In this paper Fisher et al. explore the sensitivity of CLM5's vegetation responses to assumed parameters in relation to the nitrogen cycle. Overall the paper is very thorough, clearly written and sets out to achieve its objectives. I do have a series of small suggestions (below) and a few larger suggestions. which I would outline here.

\textsf{AR: Thanks for the positive comments. We hope this version is somewhat easier to digest}

1. (Classic reviewer here - sorry) Whilst the paper does do what it said it would, I can't help thinking this paper could be more interesting if it presented the analysis in a slightly different way. Feel free to ignore all of this!

\textsf{AR: Noted...}

This paper documents three new assumptions (arguably four- section 2.1.4 as well) to CLM5 related to N; however, it doesn't give much consideration to these hypotheses, instead focussing on a broader list of impacts related to parameter choice. After reading the methods, what I was expecting (hoping) to see, was what impact these different assumptions would have on carbon fluxes. I was then expecting to see the impact of the underlying params from *these schemes (assumptions)* being probed. That isn't what we get from this paper and (my bias) I feel like this is an opportunity missed. I found it hard to interpret the impact of parameters (semi)divorced from the underlying structural assumptions. Moreover, many of the params (in my eyes) had little to do with the new code additions. For example, a lot of text is devoted to exploring the sensitivity to the Medlyn slope parameter. To me, this feels out of scope for what this paper should(could) be.

\textsf{AR: We appreciate the disconnect between the parameter choice being seemingly unrelated to the lengthy discussion of the new nitrogen cycling elements of the model. On further reflection,  we determined that our overall aim is to illustrate how the C and N responses of the model are impacted by parameter choice, given their large implications for carbon cycle feedback analysis. We think this is of slightly greater interest to a wider audience than specifically probing the differences between the CLM4.5 and CLM5 N cycle, although the two are related.  While many of the elements that affect the CO$_{2}$ and N responses are linked to the new model structure, there are several parameters that have large impacts which are not. In the new version of the manuscript, we have chosen to reduce the amount of text devoted to the specific changes to the N cycle, and instead to shift the focus slightly such that we are now looking specifically at how parameters affect the fertilization responses. Parameters of the new model components are included specifically to illustrate their impact. }

\textsf{An alternative approach would be to run the simulations with the various pieces of code turned off and returned to the default state. The problem with implementing that idea is that large changes of that type tended to modify the system state.  Therefore there are impacts on the fertilization responses solely because the initial system state was moved to conditions that were more or less light saturated, and more or less limited by water or nitrogen. (indeed, this is one of the main conclusions of the analysis). Further, with the exception of the LUNA code, technically it is quite difficult to turn off and on the FUN and FlexCN options without invoking other more poorly documented post CLM4.5 modifications, thus we have taken the approach of exploring the behaviour of the new models by modifying their major controlling parameters.}

So, my suggestion here would be to show the impact of each of the 3(4) new assumptions in relation to how nutrients are treated with CLM5 and then to explore how sensitive those results were to the underlying params. Here, my expectation was that in some cases the results would be insensitive, but in others, it is clear that the notion of a single PFT value does not make sense. As I understand it, the authors varied the gs slope term based on the param range in my 2015 paper. This gives them some erroneous results, but that wouldn't be unexpected as using the C3 grass param (a "spendy" plant) in an arid environment wouldn't make biophysical sense?

\textsf{AR: We agree with this idea wrt. the range of variation across PFTs perhaps being excessively wide. We have now modified the stomatal slope range to use the standard error reported in Table 2 of de Kauwe et al. 2015. This does notably reduce the impact of the Medlyn slope parameter on the analysis, and remove some of the more interesting feedbacks on water limitation that the wider range had illustrated, but maybe that's just how it has to be.}

In my opinion the editor shouldn't impose such a change on the authors as it is potentially a lot of work and not what they wished their paper to be about. It is just my suggestion. Equally, I would have loved to have seen those N hypotheses tested at the Duke/Oak Ridge FACE experiments, perhaps that is for another paper :P.

\textsf{AR: For the sake of full disclosure, we did actually run the model for those sites, but for reasons that are unclear, we got very low LAI predictions, and therefore extremely sensitive responses to fertilization. We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.}

2. At the very least, I think the number of params considered should be streamlined. Many feel a bit irrelevant and there is a lot of text devoted to each. Whilst I appreciate the lengths the authors have gone to explaining the impact in the results, I found myself drifting because at times it felt out of scope. I think reducing the number considered (even moving N to the supplementary) would make the paper more readable.

\textsf{AR: We agree here, and have as a result removed 5 of the 13 parameters that had limited impacts on the fertilization responses. We have also made efforts to streamline the remaining results text, and we hope the new text is not such a trial to get through (concurring that the first version was challenging in this respect). }

3. The manuscript makes a number of citations to CLM papers in review. In many places these references add little to the paper as they provide no detail and I can't look the paper up (obviously at some point soon I will be able to). My sense is that the paper would benefit from either expanding the text in some places or removing the reference entirely. For example, the Kennedy plant hydraulics paper was referenced on five occasions, that seems excessive to me given the paper is on nitrogen. This might seem like I'm being pedantic, but honestly it breaks the readability of the text.

\textsf{AR: The MS is part of a special issue, and so all the papers should become available at the same time (except the Kennedy paper which is already published). We have reduced the number of citations for the Kennedy paper down to two. The primary resource we cite is the CLM5 tech note, which is already online but will also become an appendix to the Lawrence et al. paper when these papers are all published simultaneously. Across the special issue, this is our overall strategy to documenting the entirety of the model, and we wanted to avoid duplicating large amounts of the technical text in different places.}

4. Whilst I wholeheartedly agree that assessing the model simulations against observations doesn't necessarily indicate skill, it does go a bit against the grain the way the authors have presented their results. The plots as presented are quite abstract and the authors may consider this point in revision. As a reviewer, it wouldn't bother me at all if the model seemingly performed worse with added "realism", particularly if the authors could explain why.

\textsf{AR: Fair point, but 1) you may be an atypical reviewer and 2) we still think this would distract from the main purpose of the paper, which is to document and explain the differential sensitivities of the parameter ensemble to CO$_{2}$ and N fertilization. If we added in specific site data, readers would almost certainly focus on the model fit to particular sites, which we anticipate will be explored in future literature. Nonetheless, the plots are a little abstract. We have tried to improve the presentation of the results by consolidating the number of plots, putting all of the state outputs into the same figure, adding a new figure focused on the time-dependant responses of the default model, and }

5. Finally, with the discussion I'm reading this as a non-CLM user and thinking what can I take away here? Here I struggled a little bit if I'm honest. There is some interesting text about objective calibration (great to be able to cite this :P), but again I'm wondering if this is really in the scope? Six paragraphs are devoted to apparent limitations - seems excessive, there is more to sell here than criticise? I think other modelling groups who haven't yet implemented these hypotheses in relation to the representation of the N cycle could learn a great deal from a paper like this. I'd ask the authors to reflect on what they think the messages they'd like to portray are and see if they feel those messages are in the discussion.

\textsf{AR: We concur that this aspect was difficult to ascertain from the original text, and so have re-written the discussion section with a view to including clearer digestion of the findings, and also a specific section on which parts of the analysis may and may not be applicable to models other than CLM5. We also reduced the scope of the limitations section to balance out weighting of the discussion.}

Minor stuff (Key points/Abstract/Intro/Methods): 
------------------------------------------------

- Could the second key point be written in more readable English? In the third key point N has not been defined. Also, fertilisation of what (presumably CO2? But it could be nitrogen...or anything). Finally, "other important parameters depend on the initial system state". This is hard to interpret and without context is limited as a take home point for the reader. I really suggest the authors revise the key points as I take little from them about this study.

\textsf{AR: Fair point. We have modified these points to read\\
"The Community Land Model, version 5, contains numerous modifications to representations of the vegetation carbon and nitrogen cycle.}

\textsf{The control model state, and it's responses to CO$_{2}$ and nitrogen addition are sensitive to parameter choice, but parameter responses are varied between sites that differ in their pre-fertilization water and nitrogen limitation. }

\textsf{Parameters controlling nitrogen fixation rates dominate fertilization responses. Many other fertilization responses are modified indirectly, via changes in the pre-fertilization state. "}

-In general, when the authors talk about "model state" - both in the key points and the abstract - doesn't this require some time context? Are we talking about the equilibrium state? The state at 2019?

\textsf{AR: We have changed this to 'pre-fertilization state' throughout. }

- Line 45: what would a passive nitrogen cycle look like?

\textsf{AR:  We took out "active". and changed to "included nitrogen (N) cycling and limitation of plant growth."}

- Line 46: why is it "might" the model either said that would happen or it didn't.

\textsf{AR: Changed to "In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to ... "}

- Line 52-3: Is that true? Unless that can be supported I would suggest omitting. 
\textsf{AR: We have deleted this section,}

- Line 64: It would be good not to overlook the numerous earlier studies that did actually explore these issues ... e.g. Rastetter et al. (1991) Tree Physiology 9, 101 -126.; Comins \& McMurtrie (1993) Ecological Applications 3, 666-681.; Kirscbaum et al. (1994) Plant, Cell \& Environment, 17: 1081-1099, etc.

\textsf{AR: Thanks for the references. I didn't know about those. We've deleted that section as a way to reframe the introduction to be more general and less focused on the N modifications.**}

- Line 70 onwards. Does this paragraph add anything to the introduction? If anything it is more methods anyway ...

\textsf{AR: OK, we added this part to the methods section. }

- Line 109: Is 1 really a hypothesis as stated?

\textsf{AR: We have modified the hypotheses to be "motivating questions", to re-frame the results and discussion to encompass our intended aims.}

- Line 152: I realise this is more about the LUNA model paper, but I do wonder about the assumption that day length drives variations in photosynthetic capacity, given that you already account for radiation. This just feels like a fudge for a gap in our knowledge as the mechanistic interpretation is less, clear. Anyway, irrelevant I'm sure. 
\textsf{AR: I am inclined to agree, but there's not much that can be done about this issue at this point.}

- Line 163 onwards: In the description of FUN, it would be helpful to provide a little more detail about how much of the cost is parameterised vs emergent. The text is very clear, but I still found myself wanting to know a little more. Obviously I could read the documentation!

\textsf{AR: We added a new sentence to describe the dependence of costs on the environment: \emph{The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, re-translocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (Bloom et al. [1985], Jiang et al. [2017], Terrer et al. [2018]) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation. Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. }}

- Line 243: How does the FATES model account for this issue? This seems like it might be an interesting point, but the lack of details means that this statement should either be omitted, or briefly expanded upon.

\textsf{AR: We changed this to: \emph{Newer codes within the CLM framework (e.g. the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.}}

- May I suggest (an optional) text reorganisation? Swap 2.13 and 2.12. Then 2.1.4 follows more logically after the text about FUN for the reader.

\textsf{AR: We have deleted these sections, and reorganized the grouping of the more high-level text that replaces them.}

- Line 291 - it would be hard to see a paper in review? Is it in open review? Otherwise more detail should be added, I have no issue with keeping the citation.

\textsf{AR: It is in review for the special issue at JAMES to which this paper is also submitted, and a copy should somehow be in the system to be accessed, although I don't know how referees get access to the other companion papers exactly}

- Line 303 - how is the analysis in this paper different to Wieder et al. in review? I would argue to remove this statement as it adds nothing. Similarly on line 314, reference to Kennedy. What relevance are these sentences?

\textsf{AR: Modified the Wieder reference to read: \emph{ollowing \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.}}

- Line 311 - it's 
\textsf{AR: done}

- Line 318: I think the approach used by the authors is perfectly legitimate and I see no need to expand words justifying why they did not do a global simulation. I would argue to remove these statements.

AR: \textsf{We have reduced the amount of text devoted to this idea.}

Results 
-------

- Line 423: To not look at soil outputs feels like an oversight to me. At the very least, linking the results to available plant N in the soil would be very interesting. Many of the changes have feedbacks for the soil and I think it is important to consider the responses of some of these variables too.

\textsf{AR: Think about this. **}

- Fig 2. I can't tell the difference between slatop and stem:leaf.

\textsf{AR: One is red and the other is a purple.  We have modified the color map to distinguish better between these two colours.}

- Line 509: this is another example where i feel there is too much assumed knowledge by the reader. Having not read Lawrence et al. in review, it is hard to take much from this. Given I know who Lawrence is, I just end up confused how two CLM papers would contradict each other, but lacking the details.

\textsf{AR: The point we're trying to make here is that we find that there is no saturation with increasing leafN content, which is consistent with an analysis in the Lawrence paper that finds low Vcmax compared to observations. We have changed the text to try and make it clearer that this is a point of consensus not contradiction.}

- Overall, I struggled to see what much of the description around parameter sensitivity had to do with the various nutrient assumptions. For example, the text around the Medlyn model could easily be completely omitted. This is true of a number of these parameter paragraphs. Unless there is a link to the nutrient assumptions I fail to see why these details are included? I also think if you dropped some of these irrelevant params, then the plots (fig 6-8) would be easier to interpret.

\textsf{AR: This point is well taken. As we discussed in response to a previous point, we decided to modify the focus of the manuscript to instead illustrate the parameters that are important for the C and N fertilization responses. Further, in light of the need to simplify the discussion and analysis, we have removed 5 of the 13 parameters from the initial version. The reduced range of variation for the Medlyn slope parameter further reduced the impact of that parameter on the output, and therefore less text is expended on its discussion.}

- Lines 576: These response ratios are really quite large? GPP increased by 44\% for an increase of CO2 to 550? Bigger still for NPP, e.g. 1.66 at CAX, presumably the CO2 x water interaction? Are these the responses without a nutrient cycle? I must have missed this point if so. I see that the text on line 582 would imply this to be the case. It might be worth clarifying in the results at this point.

\textsf{AR: OK, this does need clarifying. Actually, CLM5 can't be run without its N cycle. We have found in subsequent analyses that the increase in LAI with either CO2 or n can have large impacts on the response ratios, with the instantaneous response differing substantially from the longer term response on account of increases in leaf area. At Caxiuana there is a 50 \% increase in LAI on account of the CO2 fertilization, which is on top of the 33\% increase in GPP in the first year, for example. In the new version of the paper, we focus more closely on the time dependance and LAI interactions of the results. **}

- Line 701: Or that global maps are a poor way to see how comparable models and observations are ... 
\textsf{AR: Or that, yes. We don't really have enough here to say that the global maps are off though, so I'm going to desist from stating that.}

Reviewer 2

This study presents an introduction to some new modifications to the N cycle representation in CLM5 model, investigates how the model state and the responses of the system to CO2 and N addition are sensitive to parameter choices, and showed that parameters controlling N fixation dominate fertilization responses.

I find it hard to convince myself what does it mean to change parameters for each site without proper calibration. I suppose that most of these parameters are calibratable. Therefore, how these parameters change within a given climate space is somewhat pre-defined. You already suggested that these parameters are influential to plant responses to CO2 and N. Once you calibrated them against a PFT, does it mean that you have significantly reduced degree of freedom in your current testing? The authors made some good points in the discussion (L743-752), for example, calibration does not address the issue of dynamic responses. I think these points are brought in too late. I'd suggest the authors to revise their introduction with these points incorporated more explicitly.

\textsf{AR: The referee raises a number of different points here. The first point to note is that, as we discussed in L743 onwards, we did not in fact calibrate the model either at this site, nor using a global inverse calibration method, on account of the numerous difficulties encountered in that exercise.  We are pleased that the referee considers these points to be in scope and have moved the discussion of the inverse calibration exercise to the methods section (within the discussion of the perturbed parameter ensemble)}

Additionally, I totally agree that it is important to identify how model behaviors are influenced by parameter decisions. However, I don't think that this message is new, and therefore I find the results and interpretations of the results not so informative. To start, I find the section headings in the result section very confusing. For example, within section 3.1, there is only one sub-section, i.e. 3.1.1. Additionally, section 3.3 seems to be just an empty section with no real result included. The result for section 3.3 seems to be in section 3.4. Then section 3.5 is extremely long, and with only one subsection 3.5.1. I suggest the authors to revise the result section entirely, for clearer presentations, and more logical storyline.

\textsf{AR: The referee suggests that the section headings are poorly balanced, and as such we have modified the organization so that section 3.1 deals with the system state with 3.1.1. for site differences and 3.1.2 for parametric differences. Section 3.2 deals with the fertilization responses, with 3.2.1 for overall responses and 3.2.2 for parametric controls.

\textsf{AR: The referee further questions whether this message is new, and also suggests that it is important to identify how model behaviours are influenced by parameter decisions, which is the objective of this paper. We do not suggest that this particular message is new. Our model is a very heavily modified version of a very widely used land surface scheme that will likely generate many hundreds of coupled climate model runs and analysis within the context of the CMIP project - all using the default configuration. In that context, we consider it useful to specifically map out the features of the model that apply strong controls over climate responses.

We agree, however, that the results section in the first version of the manuscript was dense and difficult to follow. We have attempted to remedy this both by reducing the number of parameters considered from 13 to 8, and also by attempting to streamline the text reporting of the results to focus on the features of the output of highest importance for understanding model sensitivity. We hope this is an easier read than the prior version. 
}

Additionally, I think the second half of the abstract (L33-39) not so informative. Instead, I think the author need to be more specific about their findings. Given the increased model complexity, how are model predictions affected by these individual new modifications? Moreover, if the output space is complex, and varied by sites and variables, can the authors identify a list of weaknesses and strengths that the new modifications add to the model? For example, can the authors give more specific recommendations as to what model representations are useful under what circumstances? And, what are the areas of further investigation, both in terms of further data collection to constrain the process, or more focused model sensitivity test to identify uncertainty? I totally understand that you have a tight space in the abstract, but the current form is too general, I think.

\textsf{AR: We are not sure how it follows that the model output space being complex should lead to a list of strengths and weaknesses of the model. Indeed, assessing the validity of the process descriptions of the model is something we specifically aim to not do. We agree that the abstract is perhaps too general, **}}

Moreover, I think the authors did not present their hypotheses correctly. The 1st hypothesis presented in this study is really not a hypothesis that is testable. The impact of parameters on what? I think you need to be more specific. Right now, there is really no testable hypothesis presented. How can you prove or disprove it? Given that in the method (L307-309) you have clearly stated that you picked 13 parameters that are influential to your results, you have already "proved" your hypothesis, right? Similar, you haven't defined what a model state means, and therefore, it is my opinion that the 2nd hypothesis is not explicitly testable, either.

\textsf{AR: \textbf{Maybe we should just ditch the hypotheses?** }}

Lastly, I am not convinced what this paper differs significantly from those that they included as "under reviews". Any parameter decision should have an impact to the model predictions, otherwise you wouldn't have it as a parameter in the model. This study is a sensitivity test paper, but the message is too complicated to make any generalizable take-home messages. I think some significant re-writing is needed.

\textsf{AR: The paper differs substantially from the papers in revision, in that it explicitly tests the parameter space of the model.  The remainder of the papers assess the responses of the default configuration. Our aim here is to be transparent about the uncertainties in the model and provide guidance on the interpretation of future results.}

Specific comments

L 63 and L1082: De Kauwe. 
\textsf{AR: Fixed.}

L 88, CO2, not CO2 
\textsf{AR: Fixed.}

L 112: You haven't define what a model state you mean, and therefore, this hypothesis is not explicitly testable. 
\textsf{AR: Through the new version of the manuscript, we have used 'pre-fertilization state' to describe this condition. }

L 113 - 117: This is a very long sentence. And it's not easy to follow. Suggest rephrasing.
\textsf{AR: Changed this to \emph{ The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics. A focus on model fit at individual sites would preclude detailed consideration of how parameter variation impacts fertilization responses, as presented here}}

L268-269: Here the author refers to the full CLM5 technical report for the description of the new algorithm. I find it not satisfactory. I think the author should give all the details relevant to the result in the paper, not a section within a technical report.

\textsf{AR: We have chosen to document the model within the Lawrence et al. paper and in the technical documentation online. The Lawrence paper is within the same AGU special issue on the CLM5 to which we have submitted this manuscript, and we have previously obtained editorial support for this strategy. Indeed, the need for these papers to cross-reference one-another is the main purpose of having them in a special issue.  Further, the full description of just the new nutrient algorithm runs to many pages, which would add substantially to this already quite long manuscript, and in principle, many other parts of the model are relevant to this analysis. If we were to include these descriptions in a way that would allow a complete understanding of the technical details, they would no doubt have to reside as a supplement to this paper, and so,  this would not necessarily provide a greater utility to the reader than the two sources where are already available, Lastly, this would also constitute publishing the same material twice in different peer reviewed manuscripts.  We have nonetheless re-framed the manuscript so that it is more focused on the CO$_{2}$ and N responses, and not on the nutrient cycling updates, so we feel that this is less necessary than before.}

L 307 - 309: I am confused here. So you identified a list of 13 parameters which you think should have direct mechanistic impacts on responses to CO2 and N fertilization. Then you performed sensitivity tests around them, and highlighted that the model indeed is sensitive to these parameter decisions. Isn't the answer obvious?

\textsf{AR: Our objective here is not simply to illustrate that the parameters do have an impact (we agree that this would be self-evident) but to provide an illustrate of \emph{how} the different parameters impact the model and under what conditions.  Thus, there is value in doing the simulations subsequent to our narrowing of the parameter set to those elements we considered to have a direct impact on fertilization responses. }

L341: You say it's fine root area per unit leaf area in bold, but its calculation is the ratio of fine root biomass to leaf area? What exactly is it?

\textsf{AR: Thanks for finding this inconsistency. It is fine root mass to leaf mass. We will correct accordingly}.

L372: remove ). 
\textsf{AR: Done}

L420: Not sure how you ended up with 672 runs. 4 sites x 13 parameters x 5 values * 3 treatments = 780. Did I miss something?

\textsf{AR: We miscalculated this. It's actually 4*13*3*4 +4*3 = 636, the former term being sites * parameters * experiments * iterations and the latter being the number of control runs at the default (sites * experiments.}

L 436: "The x-axis" to the x-axis. 
\textsf{AR: Modified accordingly}

L438: Table 1, not table 1. 
\textsf{AR: Modified accordingly}

L 447: remove ). 
\textsf{AR: Done}

L458-461: Can you rephrase? I can't follow this sentence.

\textsf{AR: We have re-phrased this sentence as "Differences in the initial conditions at the sites have implications for the parametric responses. Underlying limitations by various boundary conditions (water, nutrients, light, temperature) can affect both the ability of the system to respond to fertilization and/or parameter modifications affecting other resources, as we describe further in the following sections."}

L528: Figure 1, not figure 1. 
\textsf{: Done. Thanks for catching. }

L560: the both the? 
\textsf{AR: Done. Thanks for catching.}

L 562: ))? 
\textsf{AR: Done. Thanks for catching.}

L567: Figures 6 to 8. 
\textsf{AR: We are not sure what the suggestion is here. }

L743 - 752: I think these lines should be more thoroughly discussed. This is clearly the key novelty of this research, but I am afraid that I don't see them identified, and discussed enough throughout the text. 
\textsf{AR: We have moved the key points here into the abstract, which highlights these issues earlier and makes the abstract less generic, as per an earlier point.}

L782: leaf nitrogen, not leaf Nitrogen. 
\textsf{AR: Done..}

L811: nitrogen. 
\textsf{AR: Done..}

L829: replace "(/citewieder2019" with proper text. 
\textsf{AR: Done.}

Table 1: What's default? Also, why some scenarios have the same multipliers?

\textsf{AR: ** }

\end{document}